Internal models of type theory and internal higher categories for the Univalent Foundations

### 2014-05-15 article-models-ttderived.tex
stated and proved lem:jcomp-higherjcomp

>>>>

We end this subsection with a simple application of the above lemmas that is
typical for the more general formulation of composition we have
introduced. It is a special case of a consecutive composition where we consider
a morphism $\jhom{\Gamma}{A}{B}{f}$ and a morphism
$\jhom{{{\Gamma}{B}}{Q}}{R}{S}{g}$.

\begin{lem}\label{lem:jcomp-higherjcomp}
Let $\jhom{\Gamma}{A}{B}{f}$ and $\jhom{{{\Gamma}{B}}{Q}}{R}{S}{g}$ be context
morphisms. Then the following inference rules are valid:
\begin{align*}
& \inference
  { \jfam{{{{\Gamma}{B}}{Q}}{S}}{T}
    }
  { \jfameq
      {{{{\Gamma}{A}}{\jcomp{A}{f}{Q}}}{\jcomp{A}{f}{R}}}
      {\jcomp{A}{f}{\jcomp{R}{g}{T}}}
      {\jcomp{\jcomp{A}{f}{R}}{\jcomp{A}{f}{g}}{\jcomp{A}{f}{T}}}
    }  
  \\
& \inference
  { \jfam{{{{{\Gamma}{B}}{Q}}{S}}{T}}{U}
    }
  { \jfameq
      {{{{{\Gamma}{A}}{\jcomp{A}{f}{Q}}}{\jcomp{A}{f}{R}}}{\jcomp{A}{f}{\jcomp{R}{g}{T}}}}
      {\jcomp{A}{f}{\jcomp{R}{g}{U}}}
      {\jcomp{\jcomp{A}{f}{R}}{\jcomp{A}{f}{g}}{\jcomp{A}{f}{U}}}
    }  
  \\
& \inference
  { \jterm{{{{{\Gamma}{B}}{Q}}{S}}{T}}{U}{m}
    }
  { \jtermeq
      {{{{{\Gamma}{A}}{\jcomp{A}{f}{Q}}}{\jcomp{A}{f}{R}}}{\jcomp{A}{f}{\jcomp{R}{g}{T}}}}
      {\jcomp{A}{f}{\jcomp{R}{g}{U}}}
      {\jcomp{A}{f}{\jcomp{R}{g}{m}}}
      {\jcomp{\jcomp{A}{f}{R}}{\jcomp{A}{f}{g}}{\jcomp{A}{f}{m}}}
    }  
\end{align*}
\end{lem}

\begin{proof}
We only prove the first inference rule. Let $\jfam{{{{\Gamma}{B}}{Q}}{S}}{T}$ be
a family. Then we have the judgmental equalities
\begin{align*}
\jcomp{A}{f}{\jcomp{R}{g}{T}}
& \jdeq
  \jcomp{A}{f}{\subst{g}{\ctxwk{R}{T}}}
  \tag{by definition}
  \\
& \jdeq
  \subst{\jcomp{A}{f}{g}}{\jcomp{A}{f}{\ctxwk{R}{T}}}
  \tag{by \autoref{lem:jcomp-subst}}
  \\
& \jdeq
  \subst{\jcomp{A}{f}{g}}{\ctxwk{\jcomp{A}{f}{R}}{\jcomp{A}{f}{T}}}
  \tag{by \autoref{lem:jcomp-wk}}
  \\
& \jdeq
  \jcomp{\jcomp{A}{f}{R}}{\jcomp{A}{f}{g}}{\jcomp{A}{f}{T}}.
  \tag{by definition}
\end{align*}
\end{proof}
